---
title: "BANA7051 Assignment 1"
author: "Ang Zhang"
date: "2024-08-30"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## A. Sample size of the data.

first load the data, and then take the sample size with nrow() function.

```{r}
wine <- read.csv("data/winequality-red.csv", sep = ";")
wine <- select(wine, c("fixed.acidity", "volatile.acidity", "citric.acid"))
sample_size <- nrow(wine)
print(paste('sample size is ', sample_size))
```

## B. Identify outliers

Draw a plot for each of the variables:

```{r}
for (col_name in colnames(wine))
  plot(wine[[col_name]], main = paste("distribution of", col_name), ylab = col_name)
```

One outlier can be observed in the citric.acid variable.

## C. Summarize of data.

The summary() function provides a basic summary of Min, 1st Quantile, median, third quntile, max:

```{r}
summary(wine)
```

Moreover, I would like to include standard deviation to give a little bit more insight:

```{r}
for (col_name in colnames(wine)) {
  sd = sd(wine[[col_name]])
  print(paste("standard deviation of ", col_name, ": ", round(sd, 2)))
}

```

## D. Visualize the distribution of each variable.

Draw a histogram of each variable with hist() function, and draw a density curve on top of it.

```{r}
for (col_name in colnames(wine)) {
  hist(wine[[col_name]], main = paste("histogram of", col_name), freq = F, xlab = col_name, ylab = "density")
  lines(density(wine[[col_name]]), lwd = 5, col = "blue")
}
```

## E. Any skewed distribution in D?

The fixed.acidity variable appears to be right skewed. So does the citric.acid variable.

## F. What data mining methods are used in this paper?

The author discussed linear/multiple regression (MR), neural networks (NN), and support vector machines (SVM). MR can be seen as a reduced form of NN when there's no layer of hidden node. Empirical results shows that SVM outperformed NN (and also MR) in this study case, especially for white wine.
